---
title: "General Social Survey"
author: "Sierra Lander"
date: "2024"
output: 
  html_document:
    number_sections: true
    toc: true
---


```{r include=FALSE,echo=FALSE}
require(tigerstats)
require(tidyverse)
require(foreign)
source("GrabData.R")
```




#  Question 1

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

The data in the dataframe GSSdata is from the 2018 General Social Survey. The first blocks of R-code has selected down a subset of the data to just 16 variables. It has further removed unwanted factor levels in much of the data. Examine the code in the GrabData.R file to see what it is doing. Some of the variables are categorical and others are numerical. Be sure to do a variable analysis before tackling each question.  


First question - Is opinion on the death penalty (CAPPUN) independent of gun ownership (OWNGUN)?

$H_0$: The death penalty and gun ownership are independent from eachother. 

$H_A$: The death pentaly and gun owner ship are not independent from eachother. 






## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

CAPPUN is a categorical variable
OWNGUN is a categorical variable

We use helper file cat~cat 








##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results





#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->



```{r}
barchartGC(~CAPPUN + OWNGUN,data=GSSdata)
barchartGC(~CAPPUN + OWNGUN,data=GSSdata, type="percent")

```


This graph shows that people in favor of Capital punishment are more even in gun owning and there are more people who own guns that oppose capital punishment. 



#### Numerical Descriptive Results

```{r}
table1 <- xtabs(~CAPPUN + OWNGUN, data=GSSdata)
rowPerc(table1)
colPerc(table1)
```


This table shows that people who dont own a gun and oppose capital punishment are higher that people who do own a gun. 



<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

### Inferential Results



```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```



The chi squared test shows that the p-value is 0.02022 less than 0.05 this   means we must reject HO, and conclude that CAPPUN and OWNGUN are not   independent of each other.   
  

The p-values for the fisher exact test was  0.01651, and the odds ratio was   2.271;1 this means that people who do not own a gun are more likely to   oppose capital punishment than those who do own a gun.  





<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

#  Question 2

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

Is there a difference in years of education (EDUC) between the those who did or did not vote in 2012 (VOTE12)?


$H_0$; There is no difference in years of education between those who did/did not vote. 

$H_A$: There is a difference in years of education between those who did/did not vote. 





## Methods

VOTE12 is a categorical variable 

EDUC is a numerical variable

we use helper file cat~num



<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

##Results


<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results


#### Graphical Descriptive Results


```{r}
plot(EDUC~VOTE12,data=GSSdata)
```


The graph shows that there are  more people with higher education that voted than those who did not vote.


<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

#### Numerical Descriptive Results

```{r}
favstats(EDUC~VOTE12,data=GSSdata)
```
The results above show that the average education level for people who voted is higher than those who did not ( 15.72: 13.42) 



<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

### Inferential Results

  
```{r}
t.test(EDUC~VOTE12,data=GSSdata)
```


The p-value given in the t test is 0.0000001248 this is smaller than 0.05 therefore we must reject HO.





<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

#  Question 3

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

Determine if a persons performance on the wordsum test (WORDSUM) is independent of their level of schooling (DEGREE).


## Intro


$H_0$: A persons performance rate on the wordsum test is Independent of   their level of schooling.  

$H_A$: A persons preformance level on the wordsum test is Dependent of their  schooling level.   






## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

WORDSUM is a numerical variable.   

DEGREE is a catogorical variable.   

We use helper file Num~Cat  





##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results



```{r}
plot(WORDSUM~DEGREE,data=GSSdata)
```
   
The box plot shows that the average of people with a graduates degree scored high on the Wordsum test than any other degree.   




<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

#### Numerical Descriptive Results


```{r}
favstats(WORDSUM~DEGREE,data=GSSdata)
```

The fave stats test shows that the highest median is those with a graduate degree which shows that those with a graduate degree scored higher on average on the Wordsum test than any other degree. 


<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

### Inferential Results


```{r}
anova=aov(WORDSUM~DEGREE,data=GSSdata)
summary(anova)
```
The data above shows that the p-value is 0.00000025 this number is less than 0.05 therefore we must reject the Null hypothesis. 





<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->


# Discussion and Conclusion

In conclusion for question 1. the p-values and odds ratio show that that people who do not own a gun are more likely to oppose capital punishment than those who do own a gun. Therefore we we must reject HO, and conclude that CAPPUN and OWNGUN are not   independent of each other.   

for question 2. 
The p-value given in the t test is 0.0000001248 this is smaller than 0.05 therefore we must reject HO.  results also show that the average education level for people who voted is higher than those who did not ( 15.72: 13.42) 
Therefor there is a difference in years of education between those who did/did not vote.

Question 3. shows that the p-value is 0.00000025 this number is less than 0.05 therefore we must reject the Null hypothesis.the fave stats test and box plot show that the highest median is those with a graduate degree which shows that those with a graduate degree scored higher on average on the Wordsum test than any other degree. Therefore A persons preformance level on the wordsum test is Dependent of their schooling level.   




<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

